Abstractions for Multi-Sorted Substitutions

### 2024-03-14 paper/paper.tex
Update.

  \subsection{Type Contexts}
  \begin{figure}
    \centering
    % \LibCode\KContextHelper
    \begin{minipage}{1.0\linewidth}
      \begin{minipage}{0.57\linewidth}
        \LibCode\KContexts
        \LibCode\KContextLookupI
        \LibCode\KVariableTyping
      \end{minipage}
      \begin{minipage}{0.43\linewidth}
        \LibCode\KContextExt
        \LibCode\KContextLookupII
        \vspace{10.1mm}
      \end{minipage}
    \end{minipage}
    \caption{Type Contexts}
    \label{fig:type-contexts}
  \end{figure}
  Equipped with a notion of types, we are ready to define type contexts as shown
  in Figure~\ref{fig:type-contexts}.
  \ACode{\ADef{Ctx} S} describes type contexts with bindings for variables from \ACode{S}
  as functions that map variables to types of the same sort.
  As we want our framework to support dependent types, we allow a type
  in the context to use all variables bound previously in the context.
  This is achieved with \ACode{\ADef{drop-∈} x S}, which cuts off all
  sorts from \ACode{S} that are at smaller or equal positions than a variable \ACode{x}.
  The \ACode{\ADef{\_∷ᵗ\_}} function extends a context with a new binding.
  The \ACode{\ADef{lookup}} function looks up the type of a variable in a context
  and weakens it appropriately.
  Finally, a variable typing \ACode{Γ \ADef{∋} x \ADef{∶} t} is defined in terms of \ACode{\ADef{lookup}}.

@@

  The interesting cases are:
  \begin{itemize}
  \item
    the variable typing constructor \ACode{\ACon{⊢`}}, which covers both expression- and
    type-variables, analogously to the variable term constructor;
  \item
    the typing \ACode{\ACon{⊢λ}} of a lambda abstraction, which
    weakens the codomain type \ACode{t₂}.
    This is necessary, because multi-sorted syntax allows types
    to depend on expressions, so the typing derivation for \ACode{e}
    has to account for a variable, which is not used by the type; and
  \item
    the kinding constructor \ACode{\ACon{⊢τ}} states that all types
    have kind \ACode{\ACon{★}}. This is sufficient for System F as types
    are automatically well-kinded due to intrinsic scoping.
  \end{itemize}
  To teach the framework about typing, we create a structure analogously to
  \ACode{\ACon{Syntax}}:
  \LibCode\KTyping
  The instantiation for System F is straightforward:
  \AppCode\FTypingInst

@@

  We start with \ACode{\ACon{TKit}}s, which abstract over
  variable and term \emph{typing}, and then define a
  \ACode{\ACon{TTraversal}}, which provides
  substitution-preserves-typing for all \ACode{\ACon{TKit}}s.
  \LibCode\KTKit
  The operations are completely analogous to those of a \ACode{\ACon{Kit}}, but talk
  about typing instead of terms.
  Building on the fields of a \ACode{\ACon{TKit}}, we define further operations:
  \begin{itemize}
  \item
    Using the variable/term typing, we can define a renaming/substitution typing:
    \LibCode\KMapTyping
    \ACode{ϕ \ADef{∶} Γ₁ \ADef{⇒ₖ} Γ₂} states that \ACode{ϕ} is a map that takes
    terms in \ACode{Γ₁} to terms in \ACode{Γ₂}.
  \item 
    Lifting a map preserves its typing:
    \LibCode\KLiftTyping
    % \LibCode\KLiftTypingProof
  \item 
    If a variable/term has a typing, then so does its singleton renaming/substitution:
    \LibCode\KSingleTyping
    % \LibCode\KSingleTypingProof
  \end{itemize}
  % To make it easier to explicitly pass typing kits,
  % we define the notation \ACode{ϕ \ADef{∶} Γ₁ \ADef{⇒[} TK \ADef{]} Γ₂}
  % to stand for \ACode{ϕ \ADef{∶} Γ₁ \ADef{⇒ₖ} Γ₂} of some
  % \ACode{\ACon{TKit}} \ACode{TK}.
  % \LibCode\KTypingNotation

>>>>

  \subsection{Type Contexts}
  \begin{figure}
    \centering
    % \LibCode\KContextHelper
    \begin{minipage}{1.0\linewidth}
      \begin{minipage}{0.57\linewidth}
        \LibCode\KContexts
        \LibCode\KContextLookupI
        \LibCode\KVariableTyping
      \end{minipage}
      \begin{minipage}{0.43\linewidth}
        \LibCode\KContextExt
        \LibCode\KContextLookupII
        \vspace{10.1mm}
      \end{minipage}
    \end{minipage}
    \caption{Type Contexts}
    \label{fig:type-contexts}
  \end{figure}
  Equipped with a notion of types, we are ready to define type contexts as shown
  in Figure~\ref{fig:type-contexts}.
  \ACode{\ADef{Ctx} S} describes type contexts with bindings for variables from \ACode{S}
  as functions that map variables to types of the same sort.
  As we want our framework to support dependent types, we allow a type
  in the context to use all variables bound previously in the context.
  This is achieved with \ACode{\ADef{drop-∈} x S}, which cuts off all
  sorts from \ACode{S} that are at smaller or equal positions than a variable \ACode{x}.
  The \ACode{\ADef{\_∷ᵗ\_}} function extends a context with a new binding.
  The \ACode{\ADef{lookup}} function looks up the type of a variable in a context
  and weakens it appropriately.
  Finally, a variable typing \ACode{Γ \ADef{∋} x \ADef{∶} t}
  states that \ACode{\ADef{lookup} Γ x} is \ACode{t}.

@@

  The interesting cases are:
  \begin{itemize}
  \item
    the variable rule \ACode{\ACon{⊢`}}, which covers both expression- and
    type-variables, analogously to the variable term constructor;
  \item
    the lambda rule \ACode{\ACon{⊢λ}}, which
    weakens the codomain type \ACode{t₂}.
    This is necessary, because multi-sorted syntax allows types
    to depend on expressions, so the typing derivation for \ACode{e}
    has to account for a variable, which is not used by the type; and
  \item
    the kinding rule \ACode{\ACon{⊢τ}} states that all types
    have kind \ACode{\ACon{★}}. This is sufficient for System F as types
    are automatically well-kinded due to intrinsic scoping.
  \end{itemize}
  To teach the framework about typing, we create a structure analogously to
  \ACode{\ACon{Syntax}}:
  \LibCode\KTyping
  The instantiation for System F is straightforward:
  \AppCode\FTypingInst

@@

  We start with \ACode{\ACon{TKit}}s, which abstract over
  variable and term \emph{typing}, and then define a
  \ACode{\ACon{TTraversal}}, which provides
  substitution-preserves-typing for all \ACode{\ACon{TKit}}s.
  \LibCode\KTKit
  The fields of a \ACode{\ACon{TKit}} are completely analogous to
  those of a \ACode{\ACon{Kit}}, but talk about typing instead of terms.
  Building on the fields of a \ACode{\ACon{TKit}}, we define further operations:
  \begin{itemize}
  \item
    Using the variable/term typing, we can define a renaming/substitution typing:
    \LibCode\KMapTyping
    \ACode{ϕ \ADef{∶} Γ₁ \ADef{⇒ₖ} Γ₂} states that \ACode{ϕ} is a map that takes
    terms in \ACode{Γ₁} to terms in \ACode{Γ₂}.
  \item 
    Lifting a map preserves its typing:
    \LibCode\KLiftTyping
    % \LibCode\KLiftTypingProof
  \item 
    If a variable/term has a typing, then so does its singleton renaming/substitution:
    \LibCode\KSingleTyping
    % \LibCode\KSingleTypingProof
  \end{itemize}
  % To make it easier to explicitly pass typing kits,
  % we define the notation \ACode{ϕ \ADef{∶} Γ₁ \ADef{⇒[} TK \ADef{]} Γ₂}
  % to stand for \ACode{ϕ \ADef{∶} Γ₁ \ADef{⇒ₖ} Γ₂} of some
  % \ACode{\ACon{TKit}} \ACode{TK}.
  % \LibCode\KTypingNotation
